# Missing values

In this chapter, we investigate the missing or DK(Don't know)/Refuse patterns in the following.

To this end, we use an originally coded plot missing function.

```{r, echo = FALSE}

## argument: data as a matrix or a dataframe
## output: missing value patterns plots

plot_missing <- function(data, percent = FALSE){

df_miss <- data.frame(is.na(data)) %>% 
  rownames_to_column("id")

df_miss_count <- df_miss %>% 
  group_by(across(colnames(df_miss)[-1])) %>%
  summarize(miss_count2=n()) %>%
  arrange(-miss_count2) %>%
  rowid_to_column() %>%
  pivot_longer(!c(rowid,miss_count2), names_to = "key", values_to = "missingFLG") %>%
  mutate(missing = ifelse(missingFLG=="TRUE", 1, 0)) %>%
  mutate(rowid2_fct=factor(rowid)) %>%
  rename(rowid2=rowid) %>%
  rowid_to_column() %>% 
  group_by(key) %>%
  mutate(miss_count=sum(miss_count2*missing)) %>%
  mutate(rowid=min(rowid)) %>%
  ungroup() %>%
  group_by(rowid2_fct) %>%  
  mutate(compcases = ifelse(sum(missing)==0, 1, 0)) %>%
  ungroup() %>%
  mutate(key=factor(key)) %>%
  mutate(key=fct_reorder(key,-miss_count-0.5/rowid)) %>%
  mutate(mod_missing=factor(missing-compcases)) %>%
  mutate(mod_missing=fct_reorder(mod_missing,missing-compcases)) %>%
  mutate(rowid2_fct=fct_reorder(rowid2_fct,miss_count2+0.5/rowid2))



df_miss_count_g1 <- df_miss_count %>%
  group_by(key) %>%
  summarize(miss_count=max(miss_count)) %>%
  mutate(miss_percent=10^2*miss_count/nrow(df_miss))

if(percent == TRUE){

  g1 <-   ggplot(df_miss_count_g1,aes(x = key, y = miss_percent))+
    geom_bar(stat = "identity", fill="cornflowerblue", alpha = 0.7)+
    labs(x = "", y= "% rows \n missing:")+
    scale_y_continuous(breaks=seq(0,100,25),limits=c(0,100))+
    theme_bw()+
    theme(panel.grid.major.x = element_blank(),
          panel.grid.minor = element_line(colour = "grey97", size = 0.05))+
    ggtitle("Missing value patterns")+
    theme(plot.title = element_text(size = 18))
  
} else {

  g1 <-   ggplot(df_miss_count_g1,aes(x = key, y = miss_count))+
      geom_bar(stat = "identity", fill="cornflowerblue", alpha = 0.7)+
      labs(x = "", y= "num rows \n missing:")+
      theme_bw()+
      theme(panel.grid.major.x = element_blank(),
          panel.grid.minor = element_line(colour = "grey97", size = 0.05))+
      ggtitle("Missing value patterns")+
      theme(plot.title = element_text(size = 18))

  if(max(df_miss_count_g1$miss_count)<10){
    g1 <- g1 + scale_y_continuous(breaks=seq(0,10,5),limits=c(0,10))
  }

}


df_miss_count_g2 <- df_miss_count %>%
  group_by(rowid2_fct) %>%
  summarize(miss_count2=max(miss_count2),compcases=min(compcases)) %>%
  mutate(miss_percent2=100*miss_count2/sum(miss_count2)) %>%
  mutate(compcases_fct=factor(compcases)) %>%
  mutate(compcases_fct=fct_reorder(compcases_fct,compcases))

if(percent == TRUE){

g2 <- ggplot(df_miss_count_g2,aes(x = rowid2_fct, y = miss_percent2))+
    geom_bar(stat = "identity", aes(alpha = compcases_fct), fill="cornflowerblue")+
    scale_alpha_manual(values = c(0.7,1), guide="none")+
    labs(x = "", y= "% rows")+
    scale_y_continuous(breaks=seq(0,100,25),limits=c(0,100))+
    theme_bw() +
    theme(panel.grid.major.y = element_blank(),
          panel.grid.minor = element_line(colour = "grey97", size = 0.05))+
    coord_flip()

} else {

g2 <- ggplot(df_miss_count_g2,aes(x = rowid2_fct, y = miss_count2))+
    geom_bar(stat = "identity", aes(alpha = compcases_fct), fill="cornflowerblue")+
    scale_alpha_manual(values = c(0.7,1), guide="none")+
    labs(x = "", y= "row count")+
    theme_bw() +
    theme(panel.grid.major.y = element_blank(),
          panel.grid.minor = element_line(colour = "grey97", size = 0.05))+
    coord_flip()

if(max(df_miss_count_g2$miss_count2)<10){
  g2 <- g2 + scale_y_continuous(breaks=seq(0,10,5),limits=c(0,10))
}

}


g3 <- ggplot(df_miss_count, aes(x = key,y = rowid2_fct))+
  geom_tile(aes(fill = mod_missing), color = "white")+
  labs(x = "variable", y= "missing pattern")+
  theme_classic()

if(max(df_miss_count$compcases)==1){
  x_textloc_g3 <- length(unique(df_miss_count$key))/2+0.5
  y_textloc_g3 <- unique(df_miss_count[df_miss_count$mod_missing==-1,]$rowid2_fct)
  g3 <- g3+ 
        geom_text(x=x_textloc_g3, y=y_textloc_g3, label="complete cases")+
        scale_fill_manual(values = c("grey65", "grey80", "mediumpurple1"), guide="none")
} else {
  g3 <- g3+scale_fill_manual(values = c("grey80", "mediumpurple1"), guide="none")
}


return( 
g1 + plot_spacer() + g3 + g2 +
  plot_layout(widths = c(10, 3), heights = c(3, 10))
)

}

```

Here, as a conclusion, even though there are small number of refused answers, there are no truly NA values (here truly NA values mean that we even do not know whether the respondent refused to answer, or the respondent was eager to answer but we simply failed to record it). Hence, we drop all the refused variables hereafter.

Nevertheless, we left the following explanation for illustrations that can be potentially applied to other EDVA analyses. Specifically, it includes the concept of MAR or MCAR which is crucial for giving unbiased estimate for a ground average based on data including missing values.

## Respondents’ attributes for Questions 1-3

From Q1 to Q3, we use the following variables as respondents' attributes:

- F_METRO: Whether they live in metropolitan area or not

- F_CDIVISION: Their regions

- F_AGECAT: their age categories

- F_SEX: their sex

- F_EDUCCAT: their education levels

- F_MARITAL: their marital status

- F_PARTYSUM_FINAL: their supporting parties

- F_INCOME: their income categories

Actually, there are no missing values other than "DK/Refuse" in these attributes.
The "DK/Refuse" patterns are as the following where the purple tiles represent them.
It is consistent with out intuition that the largest DK/Refuse patterns are for `F_INCOME` and `F_PARTYSUM_FINAL`.

```{r}
df55 <- df55 %>%
  mutate(F_CDIVISION=gsub(F_CDIVISION,pattern="DK/Ref",replacement = NA, ignore.case = FALSE)) %>%
  mutate(F_AGECAT=gsub(F_AGECAT,pattern="DK/REF",replacement = NA, ignore.case = FALSE)) %>%
  mutate(F_SEX=gsub(F_SEX,pattern="Refused",replacement = NA, ignore.case = FALSE)) %>%
  mutate(F_EDUCCAT=gsub(F_EDUCCAT,pattern="Don't know/Refused",replacement = NA, ignore.case = FALSE)) %>%
  mutate(F_MARITAL=gsub(F_MARITAL,pattern="Refused",replacement = NA, ignore.case = FALSE)) %>%
  mutate(F_PARTYSUM_FINAL=gsub(F_PARTYSUM_FINAL,pattern="DK/Refused/No lean",replacement = NA, ignore.case = FALSE)) %>%
  mutate(F_INCOME=gsub(F_INCOME,pattern="Refused",replacement = NA, ignore.case = FALSE))

df55_attr_to_NA <- df55 %>% 
  dplyr::select(F_METRO,
  F_CDIVISION,
  F_AGECAT,
  F_SEX,
  F_EDUCCAT,
  F_MARITAL,
  F_PARTYSUM_FINAL,
  F_INCOME)

names(df55_attr_to_NA) <- substring(names(df55_attr_to_NA), 3)
names(df55_attr_to_NA) <- substring(names(df55_attr_to_NA), 1,5)
df55_attr_to_NA %>% plot_missing(percent= TRUE)

df55_attr_to_NA <- df55_attr_to_NA %>% mutate(CDIVI = fct_relevel(CDIVI,
                                   "Pacific",
                                   "Mountain",
                                   "West North Central",
                                   "West South Central",
                                   "East North Central",
                                   "East South Central",
                                   "New England",
                                   "Mid-Atlantic",
                                   "South Atlantic")) %>%
  mutate(EDUCC = fct_relevel(EDUCC,
                                   "H.S. graduate or less",
                                   "Some College", 
                                   "College graduate+")) %>%
  mutate(INCOM = fct_relevel(INCOM,
                                   "Less than $10,000",
                                   "$10,000 to less than $20,000", 
                                   "$20,000 to less than $30,000", 
                                   "$30,000 to less than $40,000", 
                                   "$40,000 to less than $50,000",
                                   "$50,000 to less than $75,000",
                                   "$75,000 to less than $100,000",
                                   "$100,000 to less than $150,000",
                                   "$150,000 or more")) %>%
  mutate(PARTY = fct_recode(PARTY, Demo = "Dem/Lean Dem", Repub = "Rep/Lean Rep")) %>%
  mutate(INCOM = fct_recode(INCOM, 
                                   "< $10K" = "Less than $10,000",
                                   "$10K-20K" = "$10,000 to less than $20,000", 
                                   "$20K-30K" = "$20,000 to less than $30,000", 
                                   "$30K-40K" = "$30,000 to less than $40,000", 
                                   "$40K-50K" = "$40,000 to less than $50,000",
                                   "$50K-75K" = "$50,000 to less than $75,000",
                                   "$75K-100K" = "$75,000 to less than $100,000",
                                   "$100K-150K" = "$100,000 to less than $150,000",
                                   ">= $150K" = "$150,000 or more"))
```

Note that there are no large refused patterns which might affect our analyses in any variables.
If these DK/Refuse patterns were large, we would have to check whether the answers for questionnaires of interest are not severely unbiased for the DK/Refuse patterns.

## Question 1

For Q1, we use the the questions below:

EN7_W55: How much do you think human activity, such as the burning of fossil fuels, contributes to global climate change?

EN8_W55: How much do you think natural patterns in the Earth’s environment contribute to global climate change?

Similarly to the attribute information, there are no missing values but there are "refuse" answers.
The refusal patterns for these questions are presented below.
Note that the explanatory variable are written in uppercase letter, while the explained ones in lowercase.

```{r}
df55_Q1_to_NA <- df55 %>% 
  dplyr::select(
  EN7_W55,
  EN8_W55,
  ) %>%
  mutate(EN7_W55=gsub(EN7_W55,pattern="Refused",replacement = NA, ignore.case = FALSE)) %>%
  mutate(EN8_W55=gsub(EN8_W55,pattern="Refused",replacement = NA, ignore.case = FALSE))

names(df55_Q1_to_NA) <- c("human","natural")

df55_Q1_to_NA <- cbind(df55_Q1_to_NA,df55_attr_to_NA)

df55_Q1_to_NA %>% plot_missing(percent= TRUE)
df55_Q1_to_NA %>% plot_missing()
```

Like the one for respondents' attributes, there are no large refusal patterns.
Also, there are no prominent refusal correlations that we should pay attention to, since the largest refusal patterns have only one refused item.

Hence, we drop all the refused variables (now having been replaced into "NA") hereafter.

```{r }
df55_Q1_to_NA <- df55_Q1_to_NA %>%
  mutate(human = fct_relevel(human,
                                   "A great deal",
                                   "Some",
                                   "Not too much",
                                   "Not at all")) %>%
  mutate(natural = fct_relevel(natural,
                                   "A great deal",
                                   "Some",
                                   "Not too much",
                                   "Not at all"))

findf55_Q1_to_NA <- df55_Q1_to_NA %>% na.omit
```



## Question 2

For CLIM9F1 (for Form 1) and F2 (for Form 2), the questions are as below:

- How much, if at all, do you think global climate change is currently affecting…

  - CLIM9F1_a_W55: The United States (1 A great deal, 2 Some, 3 Not too much, 4 Not at all)

  - CLIM9F1_b_W55: Your local community (1 A great deal, 2 Some, 3 Not too much, 4 Not at all)

  - CLIM9F2_W55: How much, if at all, do you think global climate change is currently affecting your local community?
  
Questions from CLIM10_a_W55 to CLIM10_g_W55 are asked for respondents who answered 1 or 2 for CLIM9F1_b_W55 or CLIM9F2_W55.
  
- The alternatives are either 1 or 2: 1 Yes, a major way,  2 No, not a major way

  - CLIM10_a_W55: Droughts or water shortages

  - CLIM10_b_W55: Long periods of unusually hot weather

  - CLIM10_c_W55: Severe weather, like floods or intense storms

  - CLIM10_d_W55: Rising sea levels that erode beaches and shore lines

  - CLIM10_e_W55: Damage to forests and plant life

  - CLIM10_f_W55: Harm to animal wildlife and their habitats

  - CLIM10_g_W55: More frequent wildfires

One can see from below that in Form 1, CLIM9F1_a_W55, CLIM9F1_b_W55 are not missing,
Likewise, CLIM9F2_W55 is not missing for Form 2.



```{r}
df55_Q2_to_NA <- df55 %>% 
  dplyr::select(
  CLIM9F1_a_W55,
  CLIM9F1_b_W55,
  CLIM9F2_W55,
  CLIM10_a_W55,
  CLIM10_b_W55,
  CLIM10_c_W55,
  CLIM10_d_W55,
  CLIM10_e_W55,
  CLIM10_f_W55,
  CLIM10_g_W55
  ) 

names(df55_Q2_to_NA) <- c("usa","comm","comm2","water","hot","severe","sea","forest","animal","wildfire")

df55_Q2_to_NA %>% plot_missing()
```

Also, it is confirmed from below that the questions CLIM10_X_W55 for X=a,...,g are missing only if CLIM9F1_b_W55 or CLIM9F2_W55 is answered by 3 Not too much, 4 Not at all, or "Refused."
Therefore, the only essential missing values of CLIM10_X_W55 is the one when `comm` or `comm2` is "Refused" or when they are "Refused" by themselves.

```{r}
df55_Q2_to_NA %>% filter(is.na(water)==TRUE) %>% group_by(comm,comm2,water) %>%summarise(n())
```

We then investigate the refusal patterns here.
Although it is seen that the number of missing rows is small enough to dismiss,
let's discuss in more detail.

From the refusal patterns plot below, the most frequent refusal pattern is refusals for all variables.
It is suspected that the reason for this is that answering them was felt like time-consuming and bothering because of (1) or (2) below:

- (1) their personality (not relating to their answers)

- (2) the respondents were not sure about or did not know a concrete impact of climate change on any of the items.

In the case of (1), the missing pattern is probably "missing completely at random (MCAR)" or "missing at random (MAR)," since in this case their personality is independent of the answers for CLIM10_X_W55 (impact on one's community).

For the other top refusal patterns where only one variable is refused, it is more strongly suspected that the respondents did not know whether there were an impact on the particular item.

Hence, hereafter we can rationally assume either (1) the missing pattern is either MCAR or MAR or (2) the refused answers for these questions imply "2 No, not a major way."

In the case of (1), it is known that the relationship of explanatory variables and explained one (more precisely, the least squared estimator of the slope for a true linear model) can be inferred unbiasedly. Similar argument holds at least qualitatively for discrete explanatory variables.

In the case of (2), the missing values are no longer missing (they are "2 No").

Therefore, it suffices to simply compute the proportion of "1 Yes, a major way" to all non-missing values and analyse the difference in the proportions between supporting parties to answer Q2.

```{r}
df55_Q2_to_REF <- as.matrix(df55_Q2_to_NA)

df55_Q2_to_REF <- ifelse(is.na(df55_Q2_to_REF),"not missing",df55_Q2_to_REF)

for(i in 1:nrow(df55_Q2_to_REF)){
  for(j in 3:ncol(df55_Q2_to_REF)){
    if(df55_Q2_to_REF[i,2]=="Refused" || df55_Q2_to_REF[i,3]=="Refused"){
      df55_Q2_to_REF[i,j] <- "Refused"
    }
  }
}
  

df55_Q2_to_REF <- df55_Q2_to_REF %>%
  gsub(pattern = "Refused", replacement = NA, ignore.case = FALSE) %>%
  as.data.frame()



df55_Q2_to_REF %>% plot_missing(percent= TRUE)
df55_Q2_to_REF %>% plot_missing()
```

We therefore will drop all refused variables henceforth with some factor relabeling.

```{r}
subdf55_Q2 <- as.matrix(df55_Q2_to_NA)

subdf55_Q2 <- subdf55_Q2 %>%
  gsub(pattern = "Refused", replacement = NA, ignore.case = FALSE) %>%
  as.data.frame()

subdf55_Q2 <- subdf55_Q2 %>%
  mutate(usa = fct_relevel(usa,
                                   "A great deal",
                                   "Some",
                                   "Not too much",
                                   "Not at all")) %>%
  mutate(comm = fct_relevel(comm,
                                   "A great deal",
                                   "Some",
                                   "Not too much",
                                   "Not at all")) %>%
  mutate(comm2 = fct_relevel(comm2,
                                   "A great deal",
                                   "Some",
                                   "Not too much",
                                   "Not at all")) %>%
  mutate(water = fct_relevel(water,
                                   "Yes, a major way",
                                   "No, not a major way")) %>%
  mutate(hot = fct_relevel(hot,
                                   "Yes, a major way",
                                   "No, not a major way")) %>%
  mutate(severe = fct_relevel(severe,
                                   "Yes, a major way",
                                   "No, not a major way")) %>%  
  mutate(sea = fct_relevel(sea,
                                   "Yes, a major way",
                                   "No, not a major way")) %>%  
  mutate(forest = fct_relevel(forest,
                                   "Yes, a major way",
                                   "No, not a major way")) %>%    
  mutate(animal = fct_relevel(animal,
                                   "Yes, a major way",
                                   "No, not a major way")) %>%    
  mutate(wildfire = fct_relevel(wildfire,
                                   "Yes, a major way",
                                   "No, not a major way"))

findf55_Q2 <- cbind(subdf55_Q2,df55_attr_to_NA)
```



## Question 3

For Q3, we use the questions below about agreeing or disagreeing with more nuclear power plants to generate electricity.

- ENV2_a_W55: More offshore oil and gas drilling in U.S. waters

- ENV2_b_W55: More nuclear power plants to generate electricity

- ENV2_c_W55: More coal mining

- ENV2_d_W55: More solar panel “farms”

- ENV2_e_W55: More hydraulic fracturing, sometimes called “fracking,” for oil and natural gas

- ENV2_f_W55: More wind turbine “farms”

Similarly to the attribute information, there are no missing values but there are "refuse" answers. Those refusal patterns are presented below for questions including others such as more coal mining or more natural energy.
Note that the explanatory variable are written in uppercase letter, while the explained ones in lowercase.

```{r, fig.width=10,fig.height=15}
df55_Q3_to_NA <- df55 %>% 
  dplyr::select(
  ENV2_a_W55,
  ENV2_b_W55,
  ENV2_c_W55,
  ENV2_d_W55,
  ENV2_e_W55,
  ENV2_f_W55,) %>%
  mutate(ENV2_a_W55=gsub(ENV2_a_W55,pattern="Refused",replacement = NA, ignore.case = FALSE)) %>%
  mutate(ENV2_b_W55=gsub(ENV2_b_W55,pattern="Refused",replacement = NA, ignore.case = FALSE)) %>% 
  mutate(ENV2_c_W55=gsub(ENV2_c_W55,pattern="Refused",replacement = NA, ignore.case = FALSE)) %>%
  mutate(ENV2_d_W55=gsub(ENV2_d_W55,pattern="Refused",replacement = NA, ignore.case = FALSE)) %>%
  mutate(ENV2_e_W55=gsub(ENV2_e_W55,pattern="Refused",replacement = NA, ignore.case = FALSE)) %>%
  mutate(ENV2_f_W55=gsub(ENV2_f_W55,pattern="Refused",replacement = NA, ignore.case = FALSE))

names(df55_Q3_to_NA) <- c("oilgas","nuclear","coal","solar","fracking","wind")

df55_Q3_to_NA <- cbind(df55_Q3_to_NA,df55_attr_to_NA)

df55_Q3_to_NA %>% plot_missing(percent= TRUE)
df55_Q3_to_NA %>% plot_missing()
```

Like the one for respondents' attributes, there are no large refusal patterns.
It is interesting that the most major refusal pattern is not the one in which all of them are refused, but only one of them is refused.
Combining with the fact that the user interface of these questions in the questionnaire displays only "favor" or "oppose" options, it implies that the "refusal" means "Do not have opinion" rather than "Do not want to answer" or "refuse to answer (due to some personal reason)."
Hence, even if the proportion of refusal is larger, these refusal pattern will not make the aggregated proportion biased at least severely.

Although it is of interest to investigate whether these refusal patterns are correlated with some attributes of the respondents, the small number of the refusal sample size makes it difficult to analyse.

Also, there are no prominent refusal correlations that we should pay attention to, since the largest refusal patterns have only one refused item.

Hence, we drop all the refused variables (now having been replaced into "NA") hereafter.

```{r }
df55_Q3_to_NA <- df55_Q3_to_NA %>%
  mutate(oilgas = fct_relevel(oilgas,
                                   "Oppose",
                                   "Favor")) %>%
  mutate(nuclear = fct_relevel(nuclear,
                                   "Oppose",
                                   "Favor")) %>%
  mutate(coal = fct_relevel(coal,
                                   "Oppose",
                                   "Favor")) %>%
   mutate(solar = fct_relevel(solar,
                                   "Oppose",
                                   "Favor")) %>%
   mutate(fracking = fct_relevel(fracking,
                                   "Oppose",
                                   "Favor")) %>%
   mutate(wind = fct_relevel(wind,
                                   "Oppose",
                                   "Favor"))
  
findf55_Q3_to_NA <- df55_Q3_to_NA %>% na.omit
```
