--- 
title: "Political views and their attributes"
author: "Zhenyu Yuan and Yuki Ikeda"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---
--- 
title: "Political views and their attributes"
author: "Zhenyu Yuan and Yuki Ikeda"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

# Introduction

We want to deepen the insight given in the report by the Pew Research Center through a series of EDVA analyses:

https://www.pewresearch.org/science/2019/11/25/u-s-public-views-on-climate-and-energy/

1. Although the survey pointed out supported parties as a relevant factor, are educations or incomes also related to the opinion toward climate change issues (ex. whether the climate change is due to human activities or just natural patterns in global climate)? We address this issue upon trying to control other relevant effects such as supported parties.

2. From the survey, it is seen that Democrats are more likely to say that the federal government is doing too little for key aspects of the environment, and Democrats are more likely to think the climate change is due to human activities. With respect to this, are Democrats (Republicans) also more (less) recognizing that the climate changes are affecting their local communities (no matter what the reason of the change is) in the first place? We address this issue upon trying to control other relevant effects such as regions.

3. From the survey, it is seen that majority of people think that they should decrease the oil drilling or coal mining, and they should increase natural electricity generations such as one by wind or land heat. However, their opinion is separate with respect to whether they should increase the new clear power plants (Yes:49% vs No:49%). Here, how do factors (sex, age, party, income, regions,…) relate to the choice Yes and No.

4. The survey also includes some questions about the respondents’ personality such as being organized or intellectual. How is personality affecting one’s attitude towards climate changes and willingness to take actions to reduce climate change?

5. Has the public’s attitude towards climate change evolved over time? Does this change correlate with the change in election outcomes? Is it different across regions and groups?




<!--chapter:end:index.Rmd-->


# Data sources

Placeholder


## American Trends Panel (ATP) by the Pew Research Center
## US Election Data
## How to answer the questions above using the data

<!--chapter:end:02-data.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data transformation

In this chapter, we briefly explain the data transformation process for exploratory analyses with R.

Since, the data is already well-formatted as .sav files, we only imported them into dataframes and only rendered some factor re-labeings or re-namings after analyzing the missing value patterns.

```{r}
df55 <- foreign::read.spss("./data/W55_Oct19/ATP W55.sav", to.data.frame=TRUE)
```

Note that our data have weights which correct the potential bias of the attributes of the respondents of the questionnaires based on some other referenced large data. All the results in the existing report is based on those weights. Though strictly speaking, we should also consider those weights when we aggregate samples into graphs, for the following reasons, we simple aggregation (that is, just by simply counting the number of samples in the raw data, instead of aggregating the weights attached on the raw data)

  - Many EDVA packages including those we used in this analyses cannot deal with data with weights. Hence, it requires quite additional work.
  
  - Although one way to overcome this is to resample the raw data with sampling probabilities proportional to those weights.
  However, it would fairly augment the raw data already not small (about 4,000 rows with more than 10 columns) if we try to ensure that the sampling error would be small, which is therefore not realistic.
  
  - We have confirmed that there are no fatal discrepancies (that would change the conclusion) between the weighted aggregated results given in the existing report and our simply aggregated one given here.



<!--chapter:end:03-cleaning.Rmd-->


# Missing values

Placeholder


## Respondents’ attributes for Questions 1-3
## Question 1
## Question 2
## Question 3

<!--chapter:end:04-missing.Rmd-->


# Results

Placeholder


## Question 1
## Question 2
## Question 3

<!--chapter:end:05-results.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Interactive component



<!--chapter:end:06-interactive.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Conclusion

We have investigate the relationship between individual attributes and attitudes towards environmental issues.
Main conclusions are the followings:

- Age, income, educations, and sex are related with opinions on a wide range of environmental issues such as 
whether human activities contribute to global climate change or 
whether we should have more nuclear power uses.
The existence or the direction of impact can be different (asymmetric) by each parties.

- Republicans are not only likely to attribute climate change to natural phenomenon but also they are less likely to observe climate change affecting their communities in the first place.
Therefore, we should first review research regarding the existence of impact of climate change on our communities.

- 

- Environmental issues are may be (will be) highly influential to national elections in recent years (in the future).

<!--chapter:end:07-conclusion.Rmd-->

---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.13.0
  kernelspec:
    display_name: R
    language: R
    name: ir
---
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

```{r}
library(tigris)
library(tidyverse)
library(foreign)
library(ggalluvial)
library(dummies)
library(openxlsx)
library(sf)
```

```{r}
W33 = read.xlsx("../data/W33_Mar18/ATP W33 Weighted Resample.xlsx") %>% tibble()
W67 = read.xlsx("../data/W67_May20/ATP W67 Weighted Resample.xlsx") %>% tibble()
```

# Elections and opinion on climate change 

People's views on climate change and environmental problems differ by the party the support. For example, supporters of the GOP tend to think that policies aimed at reducing the effect of climate change will hurt the economy, while the Democrats tend to think they have benefits.

```{r}
W33 %>%
    filter(CLIM5_W33!="Refused" & F_PARTY_FINAL!="Refused") %>%
    mutate(CLIM5_W33 = factor(CLIM5_W33, levels=c('Help the U.S. economy',
                                                  'Make no difference for the U.S. economy',
                                                  'Hurt the U.S. economy'))) %>%
    mutate(CLIM5_W33=recode_factor(CLIM5_W33, `Help the U.S. economy`="Positive",
                                   `Make no difference for the U.S. economy`="Neutral",
                                    `Hurt the U.S. economy`="Negative")) %>%
    mutate(F_PARTY_FINAL=recode_factor(F_PARTY_FINAL, `Something else`="Else")) %>%
    ggplot(aes(y=1, axis1=CLIM5_W33, axis2=F_PARTY_FINAL, fill=F_PARTY_FINAL)) +
        geom_flow() +
        geom_stratum() + 
        geom_text(stat = "stratum", infer.label = TRUE) +
        scale_fill_manual(values=c("Democrat"="royalblue2", "Republican"="Red",
                                  "Independent"="yellow", "Something else"="grey")) +
        scale_x_discrete(limits=c("Perception of climate adaptation\n policies' effect on economy",
                                  "Political stance")) + ylab("") +
        coord_flip() + theme(legend.position="none")
```

Similar trends appear in other climate and environment related issues.

These association are on the individual level. If we aggregate them, we might be able to see a correlation with elections outcomes.

We have Pew's survey data about environment and climate change in 2018 and 2020, which are both election years (and the surveys are prior to elections). We used the House election data instead of Senate because all House seats but only one third of Senate seats are open for election every two years, which means there's extra sources of variation for Senate elections. We didn't use Presidential election because we don't have survey data in 2016.

Unfortunately, the survey data only allows for aggregation by census region (South, West, Northeast, and Midwest, each consisting of around 10 states). This means we will only have 4 data points for a year, which precludes regression and statistical tests. Nevertheless, the difference across the regions and the trends are very obvious.

```{r}
W67_env_dummies = dummy.data.frame(W67 %>% select(c(F_CREGION, ENVIR8_a_W67,ENVIR8_b_W67,ENVIR8_c_W67,
                                  ENVIR8_d_W67, ENVIR8_e_W67)) %>% as.data.frame(),
                 names=c('ENVIR8_a_W67','ENVIR8_b_W67','ENVIR8_c_W67','ENVIR8_d_W67','ENVIR8_e_W67'),
                 sep=":") %>%
    tibble() 

W67_ENVIR8_TOO_LITTLE =
W67_env_dummies  %>%
    group_by(F_CREGION) %>%
    summarise_all(mean) %>%
    inner_join(W67_env_dummies  %>%
                    group_by(F_CREGION) %>%
                    summarise(N=n()),
              by="F_CREGION") %>%
    select(ends_with("Too little") | starts_with("F_") | starts_with("N"))

colnames(W67_ENVIR8_TOO_LITTLE) <- colnames(W67_ENVIR8_TOO_LITTLE) %>%
                                    str_replace("ENVIR8_(.)_W67:Too little", "\\1") %>%
                                    str_to_lower()
```

```{r}
W33_env_dummies = dummy.data.frame(W33 %>% select(c(F_CREGION_FINAL, ENVIR8A_W33, ENVIR8B_W33,
                                                    ENVIR8C_W33, ENVIR8D_W33, ENVIR8E_W33)) %>% as.data.frame(),
                 names=c('ENVIR8A_W33','ENVIR8B_W33','ENVIR8C_W33','ENVIR8D_W33','ENVIR8E_W33'),
                 sep=":") %>%
    tibble() 

W33_ENVIR8_TOO_LITTLE =
W33_env_dummies  %>%
    group_by(F_CREGION_FINAL) %>%
    summarise_all(mean) %>%
    inner_join(W33_env_dummies  %>%
                    group_by(F_CREGION_FINAL) %>%
                    summarise(N=n()),
              by="F_CREGION_FINAL") %>%
    select(ends_with("Too little") | starts_with("F_") | starts_with("N")) %>%
    rename(F_CREGION = F_CREGION_FINAL)

colnames(W33_ENVIR8_TOO_LITTLE) <- colnames(W33_ENVIR8_TOO_LITTLE) %>%
                                    str_replace("ENVIR8(.)_W33:Too little", "\\1") %>%
                                    str_to_lower()
```

```{r}
envir8_2018_2020 =
bind_rows(
    W33_ENVIR8_TOO_LITTLE %>% 
        mutate(year=2018) %>%
        pivot_longer(cols = !f_cregion & !year, names_to = "Question", values_to = "too little"),
    W67_ENVIR8_TOO_LITTLE %>% 
        mutate(year=2020) %>%
        pivot_longer(cols = !f_cregion & !year, names_to = "Question", values_to = "too little")
) %>% filter(Question!="n") %>% mutate(year=as.factor(year))

```

```{r}
#A nuianse here is that in Minnesota the Democratic party goes by Democratic-Farmer-Labor,
# and in North Dakota Democratic_NPL. We need to manually handle those two.

# Also, in DC, there's no Republican candidate for Representatives,
#so we need to specify values_fill=0 to avoid producing an NA which will ruin the sum function used later.

House_elec = read.csv("../data/Election_data/1976-2020-house.csv")
House_2018_2020 = House_elec %>%
    tibble() %>%
    filter(year %in% c(2018,2020)) %>%
    mutate(party = plyr::mapvalues(party,
                             c("DEMOCRATIC-FARMER-LABOR", "DEMOCRATIC-NPL", "DEMOCRATIC-NONPARTISAN LEAGUE",
                               "DEMOCRAT", "REPUBLICAN"),
                             c("Dem", "Dem", "Dem", "Dem", "Rep"))) %>%
    filter(party %in% c("Dem", "Rep")) %>%
    group_by(year, party, state_po) %>%
    summarise(votes=sum(candidatevotes)) %>%
    ungroup() %>%
    pivot_wider(names_from = "party", values_from = "votes", values_fill = 0) 
```

```{r}
ELEC_2018 = House_2018_2020 %>% filter(year==2018)
```

```{r eval=Flase}
ELEC_2018_REGION = inner_join(ELEC_2018, st_drop_geometry(states()), by=c("state_po"="STUSPS")) %>%
    group_by(REGION) %>%
    summarise(DEM=sum(Dem), REP=sum(Rep)) %>%
    mutate(pct_dem=DEM/(DEM+REP))
```

The four census regions are shown below.

```{r}
regions() %>% 
    ggplot(aes(geometry=geometry, fill=NAME)) +
        geom_sf() + 
        scale_fill_discrete(name="Census region") +
        xlim(c(-125,-65)) +
        ylim(c(20,50))
```

In both 2018 and 2020, we can see that if people think the government isn't doing enough on climate and environment issues, they tend to support the Democratic party. Exceptions are animal and water protection. It seems the two are not so important.

Note that the support for Democrats are calculated as the ratio of people who voted for a Democratic candidate in the census region. It's not the average of average in each state.

```{r}
ELEC_2020 = House_2018_2020 %>% filter(year==2020)
ELEC_2020_REGION = inner_join(ELEC_2020, st_drop_geometry(states()), by=c("state_po"="STUSPS")) %>%
    group_by(REGION) %>%
    summarise(DEM=sum(Dem), REP=sum(Rep)) %>%
    mutate(pct_dem=DEM/(DEM+REP))
```

```{r}
ENVIR8_question = list("a"="Protect air quality",
                       "b"="Protect water quality",
                       "c"="Protect animals",
                       "d"="Protect land",
                       "e"="Deal with climate change"
                      )

```

```{r}
inner_join(ELEC_2018_REGION,
        inner_join(W33_ENVIR8_TOO_LITTLE, st_drop_geometry(regions()),
           by=c("f_cregion"="NAME")),
        by=c("REGION"="REGIONCE")) %>% 
    pivot_longer(cols = c(a,b,c,d,e), names_to = "Question", values_to = "too little") %>%
    mutate(Question=sapply(Question,
                           function(i){return(ENVIR8_question[[i]])})) %>%
    ggplot(aes(x=pct_dem, y=`too little`)) +
        geom_point(aes(shape=f_cregion)) +
        stat_smooth(method="lm", se=F) +
        facet_wrap(~Question) +
        scale_y_continuous(labels = scales::percent_format(scale=100, accuracy=1)) +
        scale_x_continuous(labels = scales::percent_format(scale=100, accuracy=1)) +
        xlab("") + ylab("") +scale_shape_discrete(name="Census region") +
        ggtitle("Support for Democrats in 2018 House elections (horiz.) vs \nPercentage who thinks the federal governmentis doing too little to...")
```

```{r}
inner_join(ELEC_2020_REGION,
        inner_join(W67_ENVIR8_TOO_LITTLE, st_drop_geometry(regions()),
           by=c("f_cregion"="NAME")),
        by=c("REGION"="REGIONCE")) %>% 
    pivot_longer(cols = c(a,b,c,d,e), names_to = "Question", values_to = "too little") %>%
    mutate(Question=sapply(Question,
                           function(i){return(ENVIR8_question[[i]])})) %>%
    ggplot(aes(x=pct_dem, y=`too little`)) +
        geom_point(aes(shape=f_cregion)) +
        geom_smooth(method="lm", se=F) +
        facet_wrap(~Question) +
        scale_y_continuous(labels = scales::percent_format(scale=100, accuracy=1)) +
        scale_x_continuous(labels = scales::percent_format(scale=100, accuracy=1)) +
        scale_shape_discrete(name="Census region") +
        xlab("") + ylab("") + 
        ggtitle("Support for Democrats in 2020 House elections (horiz.) vs \nPercentage who thinks the federal governmentis doing too little to...")
```

However, when we take the difference in the two years. It's a different story. The trend has reversed. If more people fell the government is not doing enough on the issues, more people are voting for the GOP.

What's happening here? We should take note that from 2016 to 2018, the Republicans had majority in the House, while the Democrats took over the majority in 2018. Before elections, people are likely to blame the incumbent for not taking enough action.

```{r}
cregion_code = c("Midwest"='2', "Northeast"='1',"South"='3',"West"='4')
diff_envir8 = inner_join(W33_ENVIR8_TOO_LITTLE, W67_ENVIR8_TOO_LITTLE,
                   by="f_cregion", suffix=c("2018", "2020")) %>%
            mutate(diff_a=a2018-a2020,
                diff_b=b2018-b2020,
                diff_c=c2018-c2020,
                diff_d=d2018-d2020,
                diff_e=e2018-e2020) %>%
            mutate(REGION=sapply(f_cregion, function(x){return(cregion_code[[x]])}))


diff_elec = inner_join(ELEC_2018_REGION, ELEC_2020_REGION, by="REGION", suffix=c("2018", "2020")) %>%
            mutate(diff_dem_pct=pct_dem2018-pct_dem2020)

inner_join(diff_envir8, diff_elec,
        by="REGION") %>%
    select(starts_with("diff") | f_cregion) %>%
    pivot_longer(!c(diff_dem_pct, f_cregion), names_to = "Question") %>%
    mutate(Question=sapply(Question,
                           function(i){return(ENVIR8_question[[str_sub(i,-1,-1)]])})) %>%
    ggplot(aes(x=diff_dem_pct, y=value)) +
        geom_point(aes(shape=f_cregion)) +
        geom_smooth(method = "lm", se=F) +
        facet_wrap(~Question) + scale_shape_discrete(name="Census region") + xlab("") + ylab("") + 
        scale_y_continuous(labels = scales::percent_format(scale=100)) +
        scale_x_continuous(labels = scales::percent_format(scale=100), n.breaks = 2) +
        ggtitle("Change in support for Democrats in House elections (horiz.) vs \nChange in percentage who thinks the federal governmentis doing too little to...")

```

```{r}

```

<!--chapter:end:Clim_Election.Rmd-->


# People's willingness to take action to protect the environment

Placeholder



<!--chapter:end:Pers_Env.Rmd-->

